Statistical Inference of Transport Mechanisms and Long Time Scale Behavior from Time Series of Solute Trajectories in Nanostructured Membranes.

### 2020-04-10 Ben_Manuscripts/hdphmm/Draft.tex
comments.

  % BJC2: I'm not putting a lot of effort into the introduction since I'll only write
  % an abbreviated form for my dissertation. I'll work on the intro after my defense.

@@

  In our previous work, we designed two different approaches which used
  solute time series in order to parameterize stochastic models that could be used
  to project transport on much longer timescales. In our first approach we modeled 
  solute trajectories as subordinated fractional Brownian and L\'evy motion, called
  the anomalous diffusion (AD) model. We generated solute trajectories by generating
  a series of anti-correlated hops separated by random periods of entrapment drawn 
  from a power law distribution. Our second approach treated solute motion as a Markov
  state model with state-dependent dynamics, called the Markov state-dependent dynamical
  model (MSDDM). We parameterized the state transition probabilities between
  each of eight discrete states as well as the solute dynamics within each of these
  states. We generated stochastic trajectory realizations by drawing a state
  sequence based on the transition probability matrix and incorporating the state dynamics
  while solutes were trapped in each state.

@@

  We let each solute undergo 5 $\mu$s of MD simulation. We used a time step of 2 fs
  at a pressure of 1 bar and temperature of 300K controlled by the Parinello-Rahman 
  barostat and v-rescale thermostat respectively. We recorded frames every 0.5 ns.

@@

  We start by applying the IHMM algorithm to the 3D solute center-of-mass coordinate 
  trajectories transformed relative to the closest pore center. We tracked the solute's
  motion along the pore axis with the center-of-mass $z$ coordinate. Using the nearest
  pore center as the origin, we represented the radial distance of each solute's 
  center-of-mass from the pore center in 2 dimensions, $x$ and $y$ 
  %(see Figure~\ref{fig:cartesian_cylinder}). 
  By working in cartesian coordinates, we avoid
  mathematical complexity introduced by cylindrical coordinates while estimating the 
  state sequence.

% BJC6: I don't think this figure is necessary
%  \begin{figure}
%  \centering
%  %MRS3: does this preserve the z-axis symmmetry? Is it just ensuring
%  %that you can't jump that far, or is it limitinig the values that
%  %the mean can take? Apologies if it's not clear to me yet.  I'm
%  %pretty sure it's just limiting the jump (since y_t = c + f(y_t-1)),
%  %but would be good to make that explicit - ``i.e. we need to make sure it doesn't jump too far'')
%  %BJC5: preserving z-axis symmetry isn't needed yet at this stage. I use the
%  % means in z in order to distinguish states (specifically estimating the c vector).
%  % I get the radial means this way. Then just ignore z later.
%  \includegraphics[width=0.2\textwidth]{cartesian_cylinder.pdf}
%  \caption{Using the nearest pore center as the origin, we represented the solute's
%  location along the pore axis in terms of the $z$ coordinates and their radial distance
%  from the pore centers in 2 dimensions, $x$ and $y$.}\label{fig:cartesian_cylinder}
%  \end{figure}

  We applied the IHMM to each of the 24 solute trajectories independently.
  Although the IHMM is capable of identifying an infinite number of states, 
  a Dirichlet Process tends to exhibit a ``rich get richer" effect, favoring
  a fewer number of states. By applying the algorithm to each trajectory 
  independently, we reduce the possibility of lumping together multiple 
  similar states which we would prefer to stay separated before clustering.
  Note that the states identified by the IHMM are heavily influenced by 
  the Gaussian prior placed on $\mathbf{c}$ in Equation~\ref{eqn:var}. 
  In Section~\ref{S-section:prior_guesses} of the Supporting Information we outline our method for 
  choosing reliable prior parameters. The entries of $A$ and $\Sigma$ do 
  not vary over a wide range, so the final parameters were relatively insensitive
  to their priors.

@@

  We reparameterized the time series, preserving the state sequence, in terms
  of the radial and axial coordinates ($r$, $z$) because $x$ and $y$ aren't 
  meaningful alone due to the system's radial symmetry. We converted the $x$ and 
  $y$ center-of-mass coordinates to $r$. We fixed the state sequence from the 
  3D parmeterization and applied the inference component of the IHMM procedure
  to the cylindrical trajectories in order to estimate the VAR(1) parameters in
  terms of $r$ and $z$. Since we fixed the state sequence, and we are using conjugate
  priors, the VAR parameters converge very quickly (see Figure~\ref{S-fig:fixed_state_convergence}). Therefore, we only ran 
  the inference procedure for 100 iterations. 
  %However, note that the uncertainty in their
  %estimates are a function of the number of data points used in their estimate.

@@

  We reduced the parameter space used agglomerative clustering, a hierarchical
  clustering approach which uses a linkage criteria in order to successively merge
  similar clusters until a desired intracluster distance threshold or number of
  clusters is reached. We used the Ward linkage criteria, which works to minimize 
  the sum of the squared differences within all clusters.
    %MRS6: explain why Ward (make the cluster as minimally disperse?)
   	%BJC6: I'll add some plots to the SI and then write a better justification for
	% the paper. This is lower priority in mind for the thesis.
  We elected to choose the number of clusters rather than the distance threshold.
  For our data, non-parameteric methods such as Bayesian Gaussian mixture models tend
  to delocalize the clusters in parameter space (see Section~\ref{S-section:agglomerative}
  of the Supporting Information).

@@

  Choosing the number of clusters can be somewhat subjective so
  we attempted to add some structure to the selection process by following
  a set of qualitative and quantitative guidelines. We used the silhouette 
  test in order to score the quality of clustering as a function of the number 
  of clusters chosen. For our data, the silhouette test generally favors the 
  lowest number of clusters possible (see Section~\ref{S-section:nclusters} of
  the Supporting Information). However, choosing too few clusters tends to 
  result in inadequate distinguishment between visually obvious differences in 
  dynamic behavior. This results in finalized parameter sets that are averages of
  distinct behavior which presents further problems with the predictive modeling
  that we discuss later on. We aimed to maintain the highest silhouette score,
  and thus lowest number of clusters, possible while verifying that visually 
  distinct states stayed separated. Based on these guidelines, we decided to 
  group the $A$ and $\Sigma$ parameters into five clusters each and the radial
  means into three clusters. Of the 75 distinct states allowed by this formulation,
  the solutes in this study showed behavior from 32--36 distinct clusters. While
  this may seem like a high number of states, many are sampled infrequently.

@@

  We used the parameter inference component of the IHMM algorithm in order to 
  infer $A$ and $\Sigma$ of the clustered states.
   %BJC2: I'm unsure of the precise reasoning here.
  We could not simply take the mean of the clustered $A$ and $\mathbf{e}_t$ 
  parameters because it is not clear that this is a linear 
  operation for this problem. To circumvent this problem, we modified the ($r$,
  $z$) solute trajectories so that they had a mean of zero, leaving only the 
  fluctations. We did this by the mean of each same-state segment of the 
  unclustered trajectory. We used the IHMM algorithm on this modified trajectory
  to infer the clustered state parameters by fixing the clustered state sequence.

@@

  We determined the number of hydrogen bonds between each solute and the membrane
  as a function of time. Based on the geometric criteria of Luzar and Chandler, we 
  define a hydrogen bond to exist if the distance between donor, D, and acceptor, 
  A, atoms is less than 3.5\AA~and the angle formed by $D-H \cdots A$ is less than 
  30 $\degree$.~\cite{luzar_effect_1996}

@@

  We developed a way to measure the local density of the membrane at arbitrary
  points in the unit cell. We histogrammed the positions in three dimensions
  Since our system is in a monoclinic unit cell, we periodically replicated
  the system in the +/- x, y and z directions and then chose the bounds on the
  histogram in order to create a rectanglular box encompassing the unitcell with
  a 1 nm buffer between the histogram and unit cell boundaries. We then used a regular
  grid interpolator in order to allow interpolation at arbitray points within the grid.

  We quantified solute motion using the time averaged mean squared displacment (MSD)
  of the solute center of mass trajectories. The time-averaged MSD measures all observed
  displacements over time lag $\tau$:
  \begin{equation}
  	\overline{z^2(\tau)} = \dfrac{1}{T - \tau}\int_{0}^{T - \tau} (z(t + \tau) - z(t))^2 dt
  \label{eqn:tamsd}
  \end{equation}
  where T is the length of the trajectory. We reported 1$\sigma$ confidence intervals 
  based on the results of 200 bootstrap trials. For each trail, we calculated the mean of 
  $n$ random trajectories chosen with replacement from the pool of trajectories, where 
  $n$ is the number of trajectories.

@@

  We applied the IHMM to all trajectories of each of the four solutes studied.
  Based on our choice of clustering, we identified over 30 distinct dynamical
  modes for each solute. Many of these modes are sampled very infrequently and thus
  do not contribute signficantly to a general understanding of solute transport behavior.
  In this section, we explore a subset of this data, focusing on the states
  and dominant interactions most common to solute trajectories.

  \subsubsection*{The Case of Methanol}

  %MRS6: likely just the summary of the methanol behavior should go in the main paper - the rest should be SI.
  Methanol exhbits a range of dynamic behavior. In Figure~\ref{fig:mechanism_map}, we
  plot a representative 2D methanol trajectory in cylindrical coordinates. The coordinates
  are colored according to the cluster with which methanol's dynamics are most consistent,
  which implies that this particular methanol trajectory exhibits 8 distinct dynamical 
  behaviors (in order of their first appearance, they are colored dark yellow, red, light
  blue, light green, orange, blue, teal and lime green). The solute timeseries are coupled
  with color coded bars which describe the solute's local environment and physical 
  interactions that it undergoes at that trajectory frame. From the top, they describe 
  whether the solute is associated with sodium, whether the solute is hydrogen bonding
  with the monomer as well as to what part of the monomer it is hydrogen bonding, and 
  finally, the local number density surrounding the solute.

  \begin{figure}
  \centering
  \begin{subfigure}{0.75\textwidth}
  \includegraphics[width=\textwidth]{mechanism_map.png}
  \caption{}
  \end{subfigure}
  \begin{subfigure}{0.24\textwidth}
  \includegraphics[width=\textwidth]{monomer_oxygens.png}
  \caption{}
  \end{subfigure}
  \caption{(a) We can learn a significant amount of detail about solute motion by viewing 
  solute trajectories color-coded according to their clustered dynamical behavior alongside
  plots of the dominant molecular interactions and descriptions of the membrane's local 
  environment. In the plot above, we show the radial and axial coordinates of a single
  methanol molecule over the course of the equilibrated portion of its MD simulation.
  Disjoint segments of the same color imply the solute behaves similarly at different 
  points of the trajectory because their VAR parameters place them in the same cluster. 
  There are eight distinct dynamical modes exhibited by this trajectory. In order
  of their first appearance, they are colored dark yellow, red, light blue, light green,
  orange, blue, teal and lime green. On the top of the plot are bars which represent different
  physical interactions and the membrane's local environment as a function of time. The top 
  bar is colored green at time points where solutes are associated with sodium. Methanol 
  associates with sodium ions relatively infrequently. The next bar down is colored when hydrogen
  bonds exist. The colors match the enlarged oxygen atoms highlighted in (b). Blue slivers
  correspond to hydrogen bonds donated to the monomer's carboxylate head group. Green slivers
  corrrespond to  hydrogen bonds donated to the ether linkages between the head groups
  and tails. Red slivers correspond to hydrogen bonds donated to the oxygens on the ends
  of the monomer tails. Methanol exhibits significant hydrogen bonding with all parts of
  the monomer. Which part of the monomer methanol hydrogen bonds to is heavily correlated
  to the solute's $r$ coordinate. Finally, the bar with the blue gradient represents the local 
  membrane density based on the solute's ($r$, $z$) coordinate. Darker shades of blue
  correspond to higher membrane densities. The density appears to be higher when methanol
  motion is restricted (e.g. from 75 to 550 ns). It is relatively low in areas where methanol's
  position fluctuates significantly (e.g. from 550 to 700 ns). 
  }\label{fig:mechanism_map}
  \end{figure}

  One can gain useful chemical intuition just by studying plots like
  Figure~\ref{fig:mechanism_map}. It is clear that, for methanol, association 
  with sodium is a much rarer interaction than hydrogen bonding with the 
  LLC monomers. Methanol hydrogen bonds to all regions of the monomers dependent
  on its radial position. Its fluctuations tend to be smallest when hydrogen bonded
  or in areas of high local number density. For example, from approximately 75 to 
  525 ns, methanol appears to be trapped in a high density region of the tails.
  In the time that follows, it enters a region of relatively low density where 
  fluctuations are quite large. During this time period, there is intermittent, 
  short-lived hydrogen bonding with the oxygen atoms at the tail ends, as suggested
  by the frequent state changes during that time period and red slivers in the bar
  above. Signficant hops in the $z$ direction appear to be described by the dark blue
  state. After entering the dark blue state ca. 850 ns, methanol appears to jump
  signficantly in the $z$ direction from a hydrogen bonded state with the ether 
  linkages to a hydrogen bonded state with a head group in a monomer below it.

  It is clear that there is a wealth of information in these plots, but they don't give
  the complete picture needed in order to formulate LLC membrane design principles. One
  can analyze each segment sequentially and form a clear picture of this instance of 
  methanol's behavior without the need to visulize the traejectory with software like VMD.
  However, it is still possible to gain a more general understanding of methanol's average
  behavior.

@@

  Despite some similarity in their VAR parameters, the length of time spent in states 
  3--7 appears to be dependent on the solute's radial location. Based on the 
  self-transition probabilities, we estimated the average time spent by each solute in
  each state (Figure~\ref{fig:dwell_times_MET}). States 3, 5 and 7 have long dwell times
  relative to the rest because they have high self-transition probabilities. Somewhat 
  surprisingly, both states 3 and 5 correspond to solutes located close to a pore center,
  where one might expect solute motion to be less restricted. Although solutes in states
  4 and 6 are in the viscous tail region, their dwell times are only intermediate. This
  could be due to imhomogeneity in the tail region which causes transitions between many
  different trapped states but with litle net effect on solute MSD. However, there are
  instances of states within the tail region with dwell times that exceed 100--200 ns,
  as shown by our inclusion of state 7. There are no other states with dwell times longer
  than state 5 in the pore region.

  States 1 and 2 exhibit the shortest dwell times of the states studied. This is 
  consistent with the trajectory in Figure~\ref{fig:mechanism_map}. State 2 is the 
  shortest lived state, appearing to bridge other tail region states. State 1 
  appears to represent a significant hopping state. It has a high covariance and
  occurs near the pore center, where there is less obstruction to solute motion.
  It's dwell time of ca.~8 ns, suggests that solutes stay in this highly mobile 
  state for long enough for their positions to displace signficantly.

  \begin{figure}
  \centering
  \begin{subfigure}{0.6\textwidth}
  \includegraphics[width=\textwidth]{common_states_MET.pdf}
  \caption{}\label{fig:common_states_MET_lines}
  \end{subfigure}
  \begin{subfigure}{0.35\textwidth}
  % BJC: I'll bootstrap some errorbars on this by drawing the transition probabilities 
  % from a dirichlet distribution.
  \includegraphics[width=\textwidth]{A_sigma_scatter_MET.pdf}
  \caption{}\label{fig:A_sigma_scatter_MET}
  \includegraphics[width=\textwidth]{dwell_times_MET.pdf}  %BJC5: maybe better as 95 % confidence rather than mean
  \caption{}\label{fig:dwell_times_MET}
  \end{subfigure}
  \caption{We can learn the most about methanol's behavior by studying the dynamics
  of the states common to many of the methanol trajectories. In (a), we show representative
  dynamics of six states that are common to more than one third of the solute trajectories. 
  The $y$-axis labels of the $r$ coordinate plot specify the radial mean of each clustered 
  state. The percentage of trajectories in which the states appear are labeled to the right of
  the $r$ plots. States 1--5 appear in the trajectory shown in Figure~\ref{fig:mechanism_map}
  and are colored to match for easy visual comparison. All time series have a mean of zero and are
  shifted to put the mean in line with the $y$-axis labels for the purpose of visualizing them. 
  (b) We can begin to understand the solute behaviors demonstrated in (a) by understanding their VAR 
  parameters. States 1 and 2 have higher covariances in both the axial and radial directions compared
  to all other states. This should lead to signficant solute displacement. Covariance in each dimension
  appears to have a strong correlation. States 3--6 all have relatively low covariances, which may
  be characteristic of trapped states. They are primarily distinguished based on the magnitude of their 
  autoregressive coefficients, A, and radial means. Higher values of A imply motion which wanders 
  more about the mean. States 1, 3 and 5 have radial means near 1 nm or less, which suggest that they
  occur close to the pore centers. States 2, 4 and 6 occur well within the tail region of the membrane.
  (c) Using the transition probabilities (p), we estimated the average time spent within each 
  of the states using 1000 simulations of a Bernoulli process where we continued to draw until transitioning
  out of the state with probability 1 - p. Of the prevalent states, states 1 and 
  2 exhbit the shortest dwell times alongside their high covariances. States 5 and 7 have the longest dwell times in the pore and tail regions respectively.
  }\label{fig:common_states_MET}
  \end{figure}

@@

  The types of hydrogen bonding undergone by solutes is dependent on radial
  location (see Figure~\ref{fig:hbond_pichart}). States with radial means near 1 nm
  or less (states 1, 3 and 5) typcially hydrogen bond with the monomer carboxylate
  groups and ether linkages. States with radial means greater than 1 nm (states 2,
  4, 6 and 7) tend to hydrogen bond with the monomer tails. In the past we've shown
  that there is a gradual transition from the hydrophilic to hydrophobic region in
  these membranes, meaning the monomer location is not tightly bound in $r$. Therefore,
  there are instances of hydrogen bond interactions that one might consider 
  uncharacteristic based on the state's radial mean.

@@

  States describing behavior in the tail region show a direct relationship between 
  hydrogen bond lifetimes and local density. Methanol molecules which experience state 2 
  spend their time in a low density region of the tails where they can easily break free
  of hydrogen bonds. Conversely, methanol molecules trapped in state 7 are in the densest 
  region of all states studied. It is likely difficult for methanol to esacpe these high 
  density regions and its position is further stabilized via hydrogen bonding.

>>>>

  % BJC2: I'm not putting a lot of effort into the introduction since I'll only write
  % an abbreviated form for my dissertation. I'll work on the intro after my defense.  
  %MRS7: sounds good, I won't look at this today.

@@

  In our previous 
%MRS:work
study,
  we designed two different approaches which used
  solute time series in order to parameterize stochastic models that could be used
  to project transport on much longer timescales. In our first approach we modeled 
  solute trajectories as subordinated fractional Brownian and L\'evy motion, called
  the anomalous diffusion (AD) model. We generated solute trajectories by generating
  a series of anti-correlated hops separated by random periods of entrapment drawn 
  from a power law distribution. Our second approach treated solute motion as a Markov
  state model with state-dependent dynamics, called the Markov state-dependent dynamical
  model (MSDDM). We parameterized the state transition probabilities between
  each of eight discrete states as well as the solute dynamics within each of these
  states. We generated stochastic trajectory realizations by drawing a state
  sequence based on the transition probability matrix and incorporating the state dynamics
  while solutes were trapped in each state.

@@

  We let each solute undergo 5 $\mu$s of MD simulation. We used a time step of 2 fs
  at a pressure of 1 bar and temperature of 300K controlled by the Parrinello-Rahman 
  barostat~\cite{???} and velocity rescale thermostat~\cite{???} respectively. We recorded configurations every 0.5 ns.

@@

  We start by applying the IHMM algorithm to the 3D solute center-of-mass coordinate 
  trajectories transformed relative to the closest pore center. We tracked the solute's
  motion along the pore axis with the center-of-mass $z$ coordinate. Using the nearest
  pore center as the origin, we represented the radial distance of each solute's 
  center-of-mass from the pore center in 2 dimensions, $x$ and $y$ 
  %(see Figure~\ref{fig:cartesian_cylinder}). 
  By working in Cartesian coordinates, we avoid
  mathematical complexity introduced by cylindrical coordinates while estimating the 
  state sequence.

% BJC6: I don't think this figure is necessary
%  \begin{figure}
%  \centering
%  %MRS3: does this preserve the z-axis symmetry? Is it just ensuring
%  %that you can't jump that far, or is it limiting the values that
%  %the mean can take? Apologies if it's not clear to me yet.  I'm
%  %pretty sure it's just limiting the jump (since y_t = c + f(y_t-1)),
%  %but would be good to make that explicit - ``i.e. we need to make sure it doesn't jump too far'')
%  %BJC5: preserving z-axis symmetry isn't needed yet at this stage. I use the
%  % means in z in order to distinguish states (specifically estimating the c vector).
%  % I get the radial means this way. Then just ignore z later.
%  \includegraphics[width=0.2\textwidth]{cartesian_cylinder.pdf}
%  \caption{Using the nearest pore center as the origin, we represented the solute's
%  location along the pore axis in terms of the $z$ coordinates and their radial distance
%  from the pore centers in 2 dimensions, $x$ and $y$.}\label{fig:cartesian_cylinder}
%  \end{figure}

  We applied the IHMM to each of the 24 solute trajectories independently.
  Although the IHMM is capable of identifying an infinite number of states, 
  a Dirichlet process tends to exhibit a ``rich get richer" effect, favoring
  a fewer number of states. By applying the algorithm to each trajectory 
  independently, we reduce the possibility of lumping together multiple 
  similar states which we would prefer to stay separated before clustering.
  Note that the states identified by the IHMM are heavily influenced by 
  the Gaussian prior placed on $\mathbf{c}$ in Equation~\ref{eqn:var}. 
  In Section~\ref{S-section:prior_guesses} of the Supporting Information we outline our method for 
  choosing reliable prior parameters. The entries of $A$ and $\Sigma$ do 
  not vary over a wide range, so the final parameters were relatively insensitive
  to their priors.

@@

  We reparameterized the time series, preserving the state sequence, in terms
  of the radial and axial coordinates ($r$, $z$) because $x$ and $y$ aren't 
  meaningful alone due to the system's radial symmetry. We converted the $x$ and 
  $y$ center-of-mass coordinates to $r$. We fixed the state sequence from the 
  3D parameterization and applied the inference component of the IHMM procedure
  to the cylindrical trajectories in order to estimate the VAR(1) parameters in
  terms of $r$ and $z$. Since we fixed the state sequence, and we are using conjugate
  priors, the VAR parameters converge very quickly (see Figure~\ref{S-fig:fixed_state_convergence}). Therefore, we only ran 
  the inference procedure for 100 iterations. 
  %However, note that the uncertainty in their
  %estimates are a function of the number of data points used in their estimate.

@@

  We reduced the parameter space used agglomerative clustering, a hierarchical
  clustering approach which uses a linkage criteria in order to successively merge
  similar clusters until a desired intracluster distance threshold or number of
  clusters is reached. We used the Ward linkage criteria, which works to minimize 
  the sum of the squared differences within all clusters.
    %MRS6: explain why Ward (make the cluster as minimally disperse?)
   	%BJC6: I'll add some plots to the SI and then write a better justification for
	% the paper. This is lower priority in mind for the thesis.
  We elected to choose the number of clusters rather than the distance threshold.
  For our data, non-parametric methods such as Bayesian Gaussian mixture models tend
  to delocalize the clusters in parameter space (see Section~\ref{S-section:agglomerative}
  of the Supporting Information).

@@

  Choosing the number of clusters can be somewhat subjective so
  we attempted to add some structure to the selection process by following
  a set of qualitative and quantitative guidelines. We used the silhouette 
  test in order to score the quality of clustering as a function of the number 
  of clusters chosen. For our data, the silhouette test generally favors the 
  lowest number of clusters possible (see Section~\ref{S-section:nclusters} of
  the Supporting Information). However, choosing too few clusters tends to 
  not distinguish between visually obvious differences in 
  dynamic behavior. This results in finalized parameter sets that are averages of
  distinct behavior which presents further problems with the predictive modeling
  that we discuss later on. We aimed to maintain the highest silhouette score,
  and thus lowest number of clusters, possible while verifying that visually 
  distinct states stayed separated. Based on these guidelines, we decided to 
  group the $A$ and $\Sigma$ parameters into five clusters each and the radial
  means into three clusters. Of the 75 distinct states allowed by this formulation,
  the solutes in this study showed behavior from 32--36 distinct clusters. While
  this may seem like a high number of states, many are sampled infrequently.

@@

  We used the parameter inference component of the IHMM algorithm in order to 
  infer $A$ and $\Sigma$ of the clustered states.
   %BJC2: I'm unsure of the precise reasoning here.
  We could not simply take the mean of the clustered $A$ and $\mathbf{e}_t$ 
  parameters because it is not clear that this is a linear 
  operation for this problem. To circumvent this problem, we modified the ($r$,
  $z$) solute trajectories so that they had a mean of zero, leaving only the 
  fluctuations. We did this by the mean of each same-state segment of the 
  unclustered trajectory. We used the IHMM algorithm on this modified trajectory
  to infer the clustered state parameters by fixing the clustered state sequence.

@@

  We determined the number of hydrogen bonds between each solute and the membrane
  as a function of time. Based on the geometric criteria of Luzar and Chandler, we 
  define a hydrogen bond to exist if the distance between donor, D, and acceptor, 
  A, atoms is less than 3.5\AA~and the angle formed by $D-H \cdots A$ is less than 
  $30\degree$.~\cite{luzar_effect_1996}

@@

  We developed a way to measure the local density of the membrane at arbitrary
  points in the unit cell. We histogrammed the positions in three dimensions
  Since our system is in a monoclinic unit cell, we periodically replicated
  the system in the +/- x, y and z directions and then chose the bounds on the
  histogram in order to create a rectangular box encompassing the unit cell with
  a 1 nm buffer between the histogram and unit cell boundaries. We then used a regular
  grid interpolator in order to allow interpolation at arbitrary points within the grid.

  We quantified solute motion using the time averaged mean squared displacement (MSD)
  of the solute center of mass trajectories. The time-averaged MSD measures all observed
  displacements over time lag $\tau$:
  \begin{equation}
  	\overline{z^2(\tau)} = \dfrac{1}{T - \tau}\int_{0}^{T - \tau} (z(t + \tau) - z(t))^2 dt
  \label{eqn:tamsd}
  \end{equation}
  where T is the length of the trajectory. We reported 1$\sigma$ confidence intervals 
  based on the results of 200 bootstrap trials. For each trail, we calculated the mean of 
  $n$ random trajectories chosen with replacement from the pool of trajectories, where 
  $n$ is the number of trajectories.

@@

  We applied the IHMM to all trajectories of each of the four solutes studied.
  Based on our choice of clustering, we identified over 30 distinct dynamical
  modes for each solute. Many of these modes are sampled very infrequently and thus
  do not contribute significantly to a general understanding of solute transport behavior.
  In this section, we explore a subset of this data, focusing on the states
  and dominant interactions most common to solute trajectories.

  %MRS7: more descriptive title?  What about methanol?
  \subsubsection*{The Case of Methanol}

  %MRS6: likely just the summary of the methanol behavior should go in the main paper - the rest should be SI.
  Methanol exhibits a range of dynamic behavior. In Figure~\ref{fig:mechanism_map}, we
  plot a representative 2D methanol trajectory in cylindrical coordinates. The coordinates
  are colored according to the cluster with which methanol's dynamics are most consistent,
  which implies that this particular methanol trajectory exhibits 8 distinct dynamical 
  behaviors (in order of their first appearance, they are colored dark yellow, red, light
  blue, light green, orange, blue, teal and lime green). The solute time series are coupled
  with color coded bars which describe the solute's local environment and physical 
  interactions that it undergoes at that trajectory frame. From the top, they describe 
  whether the solute is associated with sodium, whether the solute is hydrogen bonding
  with the monomer as well as to what part of the monomer it is hydrogen bonding, and 
  finally, the local number density surrounding the solute.

  \begin{figure}
  \centering
  \begin{subfigure}{0.75\textwidth}
  \includegraphics[width=\textwidth]{mechanism_map.png}
  \caption{}
  \end{subfigure}
  \begin{subfigure}{0.24\textwidth}
  \includegraphics[width=\textwidth]{monomer_oxygens.png}
  \caption{}
  \end{subfigure}
  \caption{(a) We can learn a significant amount of detail about solute motion by viewing 
  solute trajectories color-coded according to their clustered dynamical behavior alongside
  plots of the dominant molecular interactions and descriptions of the membrane's local 
  environment. In the plot above, we show the radial and axial coordinates of a single
  methanol molecule over the course of the equilibrated portion of its MD simulation.
  Disjoint segments of the same color imply the solute behaves similarly at different 
  points of the trajectory because their VAR parameters place them in the same cluster. 
  There are eight distinct dynamical modes exhibited by this trajectory. In order
  of their first appearance, they are colored dark yellow, red, light blue, light green,
  orange, blue, teal and lime green. On the top of the plot are bars which represent different
  physical interactions and the membrane's local environment as a function of time. The top 
  bar is colored green at time points where solutes are associated with sodium. Methanol 
  associates with sodium ions relatively infrequently. The next bar down is colored when hydrogen
  bonds exist. The colors match the enlarged oxygen atoms highlighted in (b). Blue slivers
  correspond to hydrogen bonds donated to the monomer's carboxylate head group. Green slivers
  correspond to  hydrogen bonds donated to the ether linkages between the head groups
  and tails. Red slivers correspond to hydrogen bonds donated to the oxygens on the ends
  of the monomer tails. Methanol exhibits significant hydrogen bonding with all parts of
  the monomer. Which part of the monomer methanol hydrogen bonds to is heavily correlated
  to the solute's $r$ coordinate. Finally, the bar with the blue gradient represents the local 
  membrane density based on the solute's ($r$, $z$) coordinate. Darker shades of blue
  correspond to higher membrane densities. The density appears to be higher when methanol
  motion is restricted (e.g. from 75 to 550 ns). It is relatively low in areas where methanol's
  position fluctuates significantly (e.g. from 550 to 700 ns). 
  }\label{fig:mechanism_map}
  \end{figure}

  One can gain useful chemical intuition just by studying plots like
  Figure~\ref{fig:mechanism_map}. It is clear that, for methanol, association 
  with sodium is a much rarer interaction than hydrogen bonding with the 
  LLC monomers. Methanol hydrogen bonds to all regions of the monomers dependent
  on its radial position. Its fluctuations tend to be smallest when hydrogen bonded
  or in areas of high local number density. For example, from approximately 75 to 
  525 ns, methanol appears to be trapped in a high density region of the tails.
  In the time that follows, it enters a region of relatively low density where 
  fluctuations are quite large. During this time period, there is intermittent, 
  short-lived hydrogen bonding with the oxygen atoms at the tail ends, as suggested
  by the frequent state changes during that time period and red slivers in the bar
  above. Significant hops in the $z$ direction appear to be described by the dark blue
  state. After entering the dark blue state ca. 850 ns, methanol appears to jump
  significantly in the $z$ direction from a hydrogen bonded state with the ether 
  linkages to a hydrogen bonded state with a head group in a monomer below it.

  It is clear that there is a wealth of information in these plots, but they don't give
  the complete picture needed in order to formulate LLC membrane design principles. One
  can analyze each segment sequentially and form a clear picture of this instance of 
  methanol's behavior without the need to visualize the trajectory with software like VMD.
  However, it is still possible to gain a more general understanding of methanol's average
  behavior.

@@

  Despite some similarity in their VAR parameters, the length of time spent in states 
  3--7 appears to be dependent on the solute's radial location. Based on the 
  self-transition probabilities, we estimated the average time spent by each solute in
  each state (Figure~\ref{fig:dwell_times_MET}). States 3, 5 and 7 have long dwell times
  relative to the rest because they have high self-transition probabilities. Somewhat 
  surprisingly, both states 3 and 5 correspond to solutes located close to a pore center,
  where one might expect solute motion to be less restricted. Although solutes in states
  4 and 6 are in the viscous tail region, their dwell times are only intermediate. This
  could be due to inhomogeneity in the tail region which causes transitions between many
  different trapped states but with little net effect on solute MSD. However, there are
  instances of states within the tail region with dwell times that exceed 100--200 ns,
  as shown by our inclusion of state 7. There are no other states with dwell times longer
  than state 5 in the pore region.

  States 1 and 2 exhibit the shortest dwell times of the states studied. This is 
  consistent with the trajectory in Figure~\ref{fig:mechanism_map}. State 2 is the 
  shortest lived state, appearing to bridge other tail region states. State 1 
  appears to represent a significant hopping state. It has a high covariance and
  occurs near the pore center, where there is less obstruction to solute motion.
  It's dwell time of ca.~8 ns, suggests that solutes stay in this highly mobile 
  state for long enough for their positions to displace significantly.

  \begin{figure}
  \centering
  \begin{subfigure}{0.6\textwidth}
  \includegraphics[width=\textwidth]{common_states_MET.pdf}
  \caption{}\label{fig:common_states_MET_lines}
  \end{subfigure}
  \begin{subfigure}{0.35\textwidth}
  % BJC: I'll bootstrap some error bars on this by drawing the transition probabilities 
  % from a Dirichlet distribution.
  \includegraphics[width=\textwidth]{A_sigma_scatter_MET.pdf}
  \caption{}\label{fig:A_sigma_scatter_MET}
  \includegraphics[width=\textwidth]{dwell_times_MET.pdf}  %BJC5: maybe better as 95 % confidence rather than mean
  \caption{}\label{fig:dwell_times_MET}
  \end{subfigure}
  \caption{We can learn the most about methanol's behavior by studying the dynamics
  of the states common to many of the methanol trajectories. In (a), we show representative
  dynamics of six states that are common to more than one third of the solute trajectories. 
  The $y$-axis labels of the $r$ coordinate plot specify the radial mean of each clustered 
  state. 
  %MRS7: it's confusing to have the R not in order of r. Emphasize this more in the caption, maybe think of something a bit different for the final paper. Reads odd to have numbers not in order along an axis.
The percentage of trajectories in which the states appear are labeled to the right of
  the $r$ plots. States 1--5 appear in the trajectory shown in Figure~\ref{fig:mechanism_map}
  and are colored to match for easy visual comparison. All time series have a mean of zero and are
  shifted to put the mean in line with the $y$-axis labels for the purpose of visualizing them. 
  (b) We can begin to understand the solute behaviors demonstrated in (a) by understanding their VAR 
  parameters. States 1 and 2 have higher covariances in both the axial and radial directions compared
  to all other states. This should lead to significant solute displacement. Covariance in each dimension
  appears to have a strong correlation. States 3--6 all have relatively low covariances, which may
  be characteristic of trapped states. They are primarily distinguished based on the magnitude of their 
  autoregressive coefficients, A, and radial means. Higher values of A imply motion which wanders 
  more about the mean. States 1, 3 and 5 have radial means near 1 nm or less, which suggest that they
  occur close to the pore centers. States 2, 4 and 6 occur well within the tail region of the membrane.
  (c) Using the transition probabilities (p), we estimated the average time spent within each 
  of the states using 1000 simulations of a Bernoulli process where we continued to draw until transitioning
  out of the state with probability 1 - p. Of the prevalent states, states 1 and 
  2 exhibit the shortest dwell times alongside their high covariances. States 5 and 7 have the longest dwell times in the pore and tail regions respectively.
  }\label{fig:common_states_MET}
  \end{figure}

@@

  The types of hydrogen bonding undergone by solutes is dependent on radial
  location (see Figure~\ref{fig:hbond_pichart}). States with radial means near 1 nm
  or less (states 1, 3 and 5) typically hydrogen bond with the monomer carboxylate
  groups and ether linkages. States with radial means greater than 1 nm (states 2,
  4, 6 and 7) tend to hydrogen bond with the monomer tails. In the past we've shown
  that there is a gradual transition from the hydrophilic to hydrophobic region in
  these membranes, meaning the monomer location is not tightly bound in $r$. Therefore,
  there are instances of hydrogen bond interactions that one might consider 
  uncharacteristic based on the state's radial mean.

@@

  States describing behavior in the tail region show a direct relationship between 
  hydrogen bond lifetimes and local density. Methanol molecules which experience state 2 
  spend their time in a low density region of the tails where they can easily break free
  of hydrogen bonds. Conversely, methanol molecules trapped in state 7 are in the densest 
  region of all states studied. It is likely difficult for methanol to escape these high 
  density regions and its position is further stabilized via hydrogen bonding.
